 \documentclass[12 pt]{article}




%\usepackage{roboto-mono}
%\usepackage{fontawesome}
%\usepackage{roboto-mono}
%\usepackage[sfdefault]{roboto}
%\usepackage{lato}

%\renewcommand{\familydefault}{\sfdefault}
\usepackage{tabularx, booktabs}
\usepackage{array,multirow,graphicx}
%\newcolumntype{Y}{>{\centering\arraybackslash}X}
\usepackage{threeparttable}
\usepackage{changepage}
%\usepackage[margin=0.70in,showframe=false]{geometry}
%\usepackage{rotating}
%\usepackage{pdflscape}
%\usepackage{amsmath}
\usepackage{url}
\usepackage{longtable}
\usepackage{fancyhdr}
\usepackage{mathptmx}

\usepackage
[
        a4paper,% other options: a3paper, a5paper, etc
        left=2cm,
        right=2cm,
        top=2cm,
        bottom=2cm,
        %vmargin=2cm to make vertical margins equal to 2cm.
        %hmargin=3cm to make horizontal margins equal to 3cm.
        % use margin=3cm to make all margins  equal to 3cm.
]
{geometry}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=black,
}
 
\urlstyle{same}

\pagestyle{fancy}
\rhead{\footnotesize{Last update: \today}}




\begin{document}



\begin{longtable}{ccccc}
%\begin{table}
 %\setlength{\tabcolsep}{3pt}
%\begin{tabular}{lcccr}



\multicolumn{5}{l}{\Large{\textbf{Onur Alt{\i}nda\u{g}}}}     \\


\multicolumn{5}{l}{Assistant Professor of Economics, Bentley University}    \\
\multicolumn{5}{l}{Turkish citizen, U.S. permanent resident}    \\

\\
\multicolumn{5}{l}{\textbf{CONTACT INFORMATION}} \\[2 pt]
\multicolumn{5}{l}{AAC 181, 175 Forest St, Waltham, MA 02452.}    \\
\multicolumn{5}{l}{oaltindag@bentley.edu}   \\
\multicolumn{5}{l}{\url{http://onuraltindag.info}}   \\



\\
\multicolumn{5}{l}{\textbf{ACADEMIC POSITIONS}}  \\[2 pt]
 \multicolumn{4}{l}{2017 - present \phantom{00} Assistant Professor, Bentley University, Economics Department.}  \\
 \multicolumn{4}{l}{2016 - 2017 \phantom{0000} David E. Bell Postdoctoral Fellow, Harvard University,}  \\
  \multicolumn{4}{l}{\phantom{2016 - 2017} \phantom{0000}  T.H. Chan School of Public Health.}  \\

\\

 \multicolumn{5}{l}{\textbf{PROFESSIONAL AFFILIATIONS}}  \\[2 pt]


   \multicolumn{4}{l}{2019 - present \phantom{00}Research Fellow, Economic Research Forum.}  \\



  \multicolumn{4}{l}{2018 - present \phantom{00}Research Associate, Development Analytics.}  \\
   \multicolumn{4}{l}{2017 - present \phantom{00}Visiting scholar, Harvard Center for Population and Development Studies.} \\ 

\\


\multicolumn{5}{l}{\textbf{EDUCATION}}\\[2 pt]
  \multicolumn{4}{l}{2010\phantom{7}-\phantom{8}2016 \phantom{0000}City University of New York, Ph.D. in Economics.}     \\
    \multicolumn{4}{l}{2007\phantom{1}-\phantom{6}2008 \phantom{000..}University Paris 1 Panth\'{e}on - Sorbonne, M.A. in Economics.} \\
\multicolumn{4}{l}{2001\phantom{1}-\phantom{0}2006 \phantom{000.} Galatasaray University,  B.A. in Economics.}  \\
\\

\multicolumn{5}{l}{\textbf{PEER-REVIEWED PUBLICATIONS }} \\ 
 \multicolumn{5}{l}{\footnotesize{$\dagger$: equal co-authorship}} \\[4 pt]



   \multicolumn{5}{l}{Do Refugees Impact Voting Behavior in the Host Country? Evidence from Syrian} \\
    \multicolumn{5}{l}{Refugee Inflows in Turkey.$^{\dagger}$} \\
     \multicolumn{5}{l}{\textbf{Public Choice}, in print.} \\  
 \multicolumn{5}{l}{Altindag, O., Kaushal, N., 2020.} \\
\\

   \multicolumn{5}{l}{Can Nonexperimental Methods Provide Unbiased Estimates of a Breastfeeding} \\
  \multicolumn{5}{l}{Intervention?: A Within-Study Comparison of Peer Counseling in Oregon. } \\
    \multicolumn{5}{l}{\textbf{Evaluation Review}, 43(3–4), 152–188.} \\
 \multicolumn{5}{l}{Altindag, O., Joyce, T., and Reeder, J. A., 2019.} \\



 \\ 

  \multicolumn{5}{l}{Effects of the Affordable Care Act Medicaid Expansion on Subjective Well-Being in the US} \\
    \multicolumn{5}{l}{Adult Population, 2010–2016.$^{\dagger}$} \\
   \multicolumn{5}{l}{\textbf{American Journal of Public Health}, 109, 1236-1271.} \\
   \multicolumn{5}{l}{Kobayashi, L., Altindag, O., Truskinovsky, Y., Berkman, L., 2019.}  \\

\\
   \multicolumn{5}{l}{Judicial Bypass for Minors Seeking Abortions in Arkansas Versus Other States.} \\
    \multicolumn{5}{l}{\textbf{American Journal of Public Health} 107, 1266-1271. 
} \\
    \multicolumn{5}{l}{Altindag, O., Joyce, T., 2017.}  \\
 
\\
   \multicolumn{5}{l}{On Measuring and Reducing Selection Bias With a Quasi-Doubly Randomized} \\
      \multicolumn{5}{l}{Preference Trial.} \\
     \multicolumn{5}{l}{\textbf{Journal of Policy Analysis and Management} 36, 438-459. } \\
      \multicolumn{5}{l}{Joyce, T., Remler, D.K., Jaeger, D.A., Altindag, O., O'Connell, S.D., Crockett, S., 2017. } \\


   \multicolumn{5}{l}{Son Preference, Fertility Decline, and the Nonmissing Girls of Turkey.} \\
     \multicolumn{5}{l}{\textbf{Demography}  53, 541-566. } \\
      \multicolumn{5}{l}{Altindag, O., 2016. } \\

\\
   \multicolumn{5}{l}{Does classroom time matter?} \\
     \multicolumn{5}{l}{\textbf{Economics of Education Review} 46, 64-77. } \\
      \multicolumn{5}{l}{Joyce, T., Crockett, S., Jaeger, D.A., Altindag, O., O'Connell, S.D., 2015.} \\

\\ \multicolumn{5}{l}{Telephone Peer Counseling of Breastfeeding Among WIC Participants: A Randomized} \\
      \multicolumn{5}{l}{ Controlled Trial.} \\
     \multicolumn{5}{l}{\textbf{Pediatrics} 134, e700-e709. } \\
      \multicolumn{5}{l}{Reeder, J.A., Joyce, T., Sibley, K., Arnold, D., Altindag, O., 2014. } \\


\\


\multicolumn{5}{l}{\textbf{WORKING PAPERS}}\\[2 pt]

 \multicolumn{5}{l}{The impact of humanitarian assistance on the economic well-being of conflict-driven refugees.$^{\dagger}$} \\
 \multicolumn{5}{l}{Altindag, O., O'Connell Stephen S.D., 2020.}  \\
  \multicolumn{5}{l}{(Pre-registered, ongoing analysis)}  \\

  \\

\multicolumn{5}{l}{Targeting humanitarian aid using administrative data: model design and validation.} \\ 
  \multicolumn{5}{l}{Altindag, O., O'Connell Stephen S.D., Sasmaz A., Balcioglu Z., Jerneck M., Foong A.K., 2019.} \\
\multicolumn{5}{l}{(submitted)} \\ 
 
\\
  \multicolumn{5}{l}{Blessing or Burden? The Impact of Refugees on Businesses and the Informal Economy.$^{\dagger}$} \\

  \multicolumn{5}{l}{Altindag, O., Bakis, O., Rozo, S., 2018. (submitted)} \\
  
\\






\multicolumn{5}{l}{The Great Recession, Older Workers with Disabilities, and Implications for Retirement Security.$^{\dagger}$} \\
\multicolumn{5}{l}{Michigan Retirement Research Center Research Paper, No. 2012-277.} \\
\multicolumn{5}{l}{Altindag, O., Schmidt, L., Sevak, P. 2012. (dormant) } \\
\\


  \multicolumn{5}{l}{\textbf{GRANTS}}\\[2 pt]
 \multicolumn{5}{l}{2019 - 2020} \\  
  \multicolumn{5}{l}{Bentley University, BRC Grant.} \\
\multicolumn{5}{l}{Award amount: \$7,000, Role: PI}  \\
\\  
 \multicolumn{5}{l}{2018 - 2022} \\
 \multicolumn{5}{l}{Novo Nordisk Foundation}  \\
\multicolumn{5}{l}{The Socioeconomic Impact of Research-based Innovations and Long-term Consequences of}  \\
  \multicolumn{5}{l}{Population-based Childhood Vaccine Programs.} \\
  \multicolumn{5}{l}{with Torben Tran{\ae}s, Jane Greve, and Erdal Tekin}  \\
  \multicolumn{5}{l}{Award amount: \$1,543,500, Role: Co-PI  }  \\
  \\
  \\

 \multicolumn{5}{l}{2018 - 2019} \\
 \multicolumn{5}{l}{Bentley University, FAC Grant.} \\
\multicolumn{5}{l}{Award amount: \$2,500, Role: PI}  \\

 \multicolumn{5}{l}{Bentley University, TLN Research Seed Fund.} \\
\multicolumn{5}{l}{Award amount: \$1,500, Role: PI} \\

\\
\\
\\
\multicolumn{5}{l}{\textbf{PRESENTATIONS and SEMINARS}} \\[2 pt]

\multicolumn{5}{l}{2020 (including scheduled)} \\
\multicolumn{5}{l}{Population Association of America, (PAA 2020).} \\ 
\multicolumn{5}{l}{Economics Department Seminar, Emory University.} \\ 
\multicolumn{5}{l}{Conference on Forced Displacement, Joint Data Center on Forced Displacement in Copenhagen.}\\
\\
\multicolumn{5}{l}{2019} \\
\multicolumn{5}{l}{Research Initiative on the Economics of the Middle East (RIEME 2019), Heriot-Watt University.} \\
\multicolumn{5}{l}{Friday Lunch Seminar, Harvard Center for Population and Development Studies.} \\
\multicolumn{5}{l}{Population Association of America (PAA 2019).} \\
\multicolumn{5}{l}{International Health Economics Association (iHEA 2019).} \\
\multicolumn{5}{l}{IZA Annual Migration Meeting, UCD Geary Institute for Public Policy.} \\

\\ 

\multicolumn{5}{l}{2018} \\
\multicolumn{5}{l}{Research Initiative on the Economics of the Middle East (RIEME 2018), Paris Sorbonne University.}\\ 
     \multicolumn{5}{l}{Conference on the Impacts of Refugees in Hosting Economies,  USC Marshall School of Business.} \\
      \multicolumn{5}{l}{Social Policy and Development Seminars, Bo\u{g}azi\c{c}i University.} \\
   \multicolumn{5}{l}{Seminar, The Rockwool Foundation.}\\
         \multicolumn{5}{l}{Seminar, VIVE - The Danish Center for Social Science Research.} \\
      \multicolumn{5}{l}{Midwest Economics Association, SOLE special session.} \\
\\
  \multicolumn{5}{l}{2017}\\
      \multicolumn{5}{l}{Research Initiative on the Economics of the Middle East (RIEME 2018), Oxford University.} \\
 \multicolumn{5}{l}{Friday Lunch Seminar, Harvard Center for Population and Development Studies.}\\
  \multicolumn{5}{l}{Economic Department Seminar, Bentley University.  }\\
  \multicolumn{5}{l}{Eastern Economic Association (EEA 2017)} \\  
  \multicolumn{5}{l}{Population Association of America (PAA 2017)} \\
    \multicolumn{5}{l}{Economics Department Seminar, Ko\c{c} University} \\
    \multicolumn{5}{l}{International Health Economics Association (iHEA 2017).}\\
  

\\
   
   \multicolumn{5}{l}{2016}\\

\multicolumn{5}{l}{Conference on Human Capital, Galatasaray University.}   \\
\multicolumn{5}{l}{Economics Department Seminar, Bo\u{g}azi\c{c}i University.}   \\
\multicolumn{5}{l}{Economics Department Seminar, Sabanc{\i} University.}   \\
\multicolumn{5}{l}{Economics Department Seminar, Ko\c{c} University.}   \\
\multicolumn{5}{l}{Economics Department Seminar, Bilgi University.}   \\
\multicolumn{5}{l}{Economics Department Seminar, TOBB University.}   \\
\multicolumn{5}{l}{Economics Department Seminar, Auckland University of Technology.}   \\
\multicolumn{5}{l}{Economics Department Seminar, Bentley University.}   \\


\\
\\


 \multicolumn{5}{l}{\textbf{FELLOWSHIPS}}\\[2 pt] 
 \multicolumn{4}{l}{2016 - 2017} \\
 \multicolumn{4}{l}{David E. Bell Postdoctoral Fellowship} \\
\multicolumn{4}{l}{T.H. Chan School of Public Health, Harvard University.}   \\
\\


 \multicolumn{4}{l}{2014 - 2016} \\
 \multicolumn{4}{l}{Information Technology Fellowship} \\
\multicolumn{4}{l}{Zicklin's Online Learning and Evaluation Initiative, Baruch College, CUNY.}   \\

\\

 \multicolumn{4}{l}{2010 - 2014} \\
 \multicolumn{4}{l}{Enhanced Chancellor's Fellowship}  \\
\multicolumn{4}{l}{The Graduate Center, CUNY.}   \\

\\


 \multicolumn{5}{l}{\textbf{TEACHING EXPERIENCE}}\\[2 pt] 
  \multicolumn{4}{l}{2017 - present} \\
\multicolumn{4}{l}{Bentley University}   \\
 \multicolumn{4}{l}{Introduction to Econometrics ($\times$2)}      \\
  \multicolumn{4}{l}{Intermediate Price Theory ($\times$2)}    \\
 \multicolumn{4}{l}{Mathematical Statistics}      \\
\multicolumn{4}{l}{Principles of Microeconomics ($\times$4)} \\

\\
 \multicolumn{4}{l}{2011 - 2014} \\
\multicolumn{4}{l}{Baruch College, City University of New York}     \\

\multicolumn{4}{l}{Statistical Analysis for Economics and Finance ($\times$2)}        \\

\\

\multicolumn{5}{l}{\textbf{PROFESSIONAL EXPERIENCE}}\\[2 pt] 
\multicolumn{4}{l}{\textbf{Consulting}} \\ 
\multicolumn{4}{l}{2018-present} \\ 
\multicolumn{4}{l}{Lead Specialist/Principal Investigator}    \\
\multicolumn{4}{l}{United Nations High Commissioner for Refugees.}   \\
\multicolumn{4}{l}{World Food Organization.}   \\

\\
\multicolumn{4}{l}{\textbf{Research Assistance}} \\ 
\multicolumn{4}{l}{2012-2016} \\ 
\multicolumn{4}{l}{National Bureau of Economic Research, New York.}  \\
\\



\multicolumn{5}{l}{\textbf{SKILLS}}\\[2 pt]

 \multicolumn{4}{l}{\textbf{Language}: Turkish (native), English, French.}      \\
  \multicolumn{4}{l}{\textbf{Software}: STATA, R.}\\
 \\


\multicolumn{4}{l}{\textbf{Pre-PhD}} \\[2 pt] 
\multicolumn{4}{l}{2009-2010} \\ 
\multicolumn{4}{l}{Research Assistant, Bah\c{c}e\c{s}ehir University Center for Economic and Social Research, Istanbul.}    \\

\\
\multicolumn{4}{l}{2009-2010} \\ 
\multicolumn{4}{l}{Method and Quality Engineer, FactSet JCF Estimates, Paris.}  \\




%\end{tabular}
%\end{table}
\end{longtable}

 




\end{document}